{\bf Landslide: A New Race-Finding Tool for 15-410}

### 2015-09-30 landslide-lecture/2015/lecture.tex
Update landslide lecture to have better picture
update lecture again
Tweak phrasing in latest landslide lecture.
Update landslide lecture for fall 2015

\institute[CMU 15-410]{Carnegie Mellon University - 15-410}
\date[]{2015, February 11}

@@

\begin{frame}{Outline}
	\textbf{Theory: Seeing race conditions in a new way}
	\begin{itemize}
		\item Case study (example)
		\item Tabular execution traces
		\item The execution tree
	\end{itemize}
	{\bf Research Technique: ``Systematic testing''}
	% Say: A way of enumerating all possible interleavings of a concurrent test case.
	\begin{itemize}
		\item Preemption points
		\item Challenges and feasibility
	\end{itemize}
	{\bf Tool: Landslide}
	\begin{itemize}
		\item How it works
		\item Coping with state space explosion
		\item Automatically choosing preemption points
		\item User study (that's you!)
			% !!! SAY !!! : If you are struggling, this is not for you, etc.
	\end{itemize}
\end{frame}

@@

\begin{frame}{Testing}
	How can programmers be confident in the correctness of their code?
	\begin{itemize}
		\item Unit tests
		\begin{itemize}
			\item good for basic functionality, bad for concurrency
		\end{itemize}
		\item Stress tests
		\begin{itemize}
			\item state of the art in 15-410
		\end{itemize}
		\item Theorem proving
		\begin{itemize}
			\item heavy burden on the programmers
		\end{itemize}
		\item Releasing to paying customers and worrying about correctness later
	\end{itemize}
	\linegap

@@

\begin{frame}{Testing Mechanisms}
	\textbf{Stress testing}: \texttt{largetest}, \texttt{mandelbrot} and friends
	\begin{itemize}
		\item Attempting to exercise as many interleavings as practical
		\item Exposes race conditions at random
		\begin{itemize}
			\item ``If a preemption occurs at just the right time\ldots''
		\end{itemize}
		\item Cryptic panic messages or machine reboots
	\end{itemize}
	\linegap
	What if\ldots
	\begin{itemize}
		\item Make educated guesses about when to preempt
		\item Preempt enough times to run {\em every single} interleaving
		\item Tell the story of what {\em actually happened}.
		\item Overlook fewer bugs!
	\end{itemize}
	% Say: At the end of this lecture, I'm going to be looking for some among you...
\end{frame}

@@

\begin{frame}{About The Project}
	\textbf{About me}: 4th year graduate student, advised by Garth Gibson
	\begin{itemize}
		\item TAed 15-410 for 3 semesters during undergrad
		\item 1st graduate year was 5th year MS, rest in Ph.D. program
		\begin{itemize}
			\item \url{http://www.contrib.andrew.cmu.edu/~bblum/thesis.pdf}
		\end{itemize}
	\end{itemize}
	\pause
	\linegap

	{\bf About Landslide}
	\begin{itemize}
		\item Simics module, which traces:
			\begin{itemize}
				\item Every instruction executed
				\item Every memory access read/written
			\end{itemize}
			% FIXME for next year -- how much slowdown is there? 20x?
		\item Originally supported only P3s; can now test P2s fully-automated
		\item {\em Landslide} shows how your {\em Pebbles} programs are not stable.
	\end{itemize}
\end{frame}

@@

%\begin{frame}{Anatomy}
%	\begin{center}
%	\includegraphics[width=0.6\textwidth]{landslide.png}
%	\end{center}
%\end{frame}

@@

\begin{frame}{Preliminary Evaluation}
	We tested 22 random submitted P2s from Spring 2014.
	\begin{itemize}
		\item Test cases:
		\begin{itemize}
			\item {\tt thr\_exit\_join}
			\item {\tt paraguay}
			\item {\tt rwlock\_downgrade\_read\_test}
			\item {\tt broadcast\_test} (new)
		\end{itemize}
		\item Each test given CPU budget of 11 cores $\times$ 10 minutes.
	\end{itemize}
\end{frame}

\begin{frame}{Preliminary Results}
	88 total tests were run (9,680 total CPU-minutes)
	\linegap

	Found 7 {\bf deterministic} initialization bugs (e.g. use-after-free)
	\linegap

	56 tests ran to completion without finding bugs:
	\begin{itemize}
		\item 47 exhausted the CPU budget.
		\item 9 completed all state spaces before time ran out.
	\end{itemize}
	\linegap

	Between 20 and 33 {\bf non-deterministic} bugs were found:
	\begin{itemize}
		\item 20 deadlock races among 12 P2s
		\item 10 use-after-free races among 6 P2s
		\item 3 misc (assert fail, segfault) among 2 P2s
	\end{itemize}
\end{frame}

@@

\begin{frame}{User Study}
	\textbf{Try Landslide on your P2!}
	\begin{itemize}
		\item Bare minimum effort: No more than 1 hour
		\begin{itemize}
			\item Clone a github URL, run setup script, run, answer survey questions
			\item Landslide will automatically report test results
		\end{itemize}
		\item Full study plan: 4-8 hours of active attention
		\begin{itemize}
			\item (Estimated, including time to diagnose and fix bugs)
			\item However, many tests should run passively overnight -- start soon!
		\end{itemize}
	\end{itemize}
	\pause
	\linegap

@@

	To participate\dots
	\begin{itemize}
		\item Review this lecture and study information sheet
		\item Meet prerequisites of passing P2 tests
		\item Email me ({\tt bblum@cs.cmu.edu}) for instructions, survey
		\item Use Landslide in addition to stress tests until P2 is due!
	\end{itemize}
	\linegap

@@

\begin{frame}{More Preliminary Results}
	\textbf{Is dynamically adding ``data race'' PPs effective?}
	\begin{itemize}
		\item 3 among reported bugs (9-15\%) required data races to expose.
	\end{itemize}
	\linegap

>>>>

\institute[CMU 15-410]{Carnegie Mellon University - 15-410}
\date[]{2015, September 30}

@@

\begin{frame}{Outline}
	\textbf{Theory: Seeing race conditions in a new way}
	\begin{itemize}
		\item Case study (example)
		\item Tabular execution traces
		\item The execution tree
	\end{itemize}
	{\bf Research Technique: ``Systematic testing''}
	% Say: A way of enumerating all possible interleavings of a concurrent test case.
	\begin{itemize}
		\item Preemption points
		\item Challenges and feasibility
	\end{itemize}
	{\bf Tool: Landslide}
	\begin{itemize}
		\item How it works
		\item Automatically choosing preemption points
		\item User study (that's you!)
			% !!! SAY !!! : If you are struggling, this is not for you, etc.
	\end{itemize}
\end{frame}

@@

\begin{frame}{Testing}
	How can programmers be confident in the correctness of their code?
	\begin{itemize}
		\item Unit tests
		\begin{itemize}
			\item good for basic functionality, bad for concurrency
		\end{itemize}
		\item Stress tests
		\begin{itemize}
			\item state of the art in 15-410
		\end{itemize}
		\item Theorem proving
		\begin{itemize}
			\item heavy burden on the programmers
		\end{itemize}
		\item Releasing to paying customers and worrying about correctness later
			% Say... "and I'm hoping to prevent you from finding yourself in a future where this happens!"
	\end{itemize}
	\linegap

@@

\begin{frame}{Testing Mechanisms}
	\textbf{Stress testing}: \texttt{largetest}, \texttt{mandelbrot} and friends
	\begin{itemize}
		\item Attempting to exercise as many interleavings as practical
		\item Exposes race conditions at random
		\begin{itemize}
			\item ``If a preemption occurs at just the right time\ldots''
		\end{itemize}
		\item Cryptic panic messages when failure occurs
	\end{itemize}
	\linegap
	What if\ldots
	\begin{itemize}
		\item Make educated guesses about when to preempt
		\item Preempt enough times to run {\em every single} interleaving
		\item Tell the story of what {\em actually happened}.
		\item Overlook fewer bugs!
	\end{itemize}
	% Say: At the end of this lecture, I'm going to be looking for some among you...
\end{frame}

@@

\begin{frame}{About The Project}
	\textbf{About me}: 5th year graduate student, advised by Garth Gibson
	\begin{itemize}
		\item TAed 15-410 for 3 semesters during undergrad
		\item 1st graduate year was 5th year MS, rest in Ph.D. program
		\begin{itemize}
			\item \url{http://www.contrib.andrew.cmu.edu/~bblum/thesis.pdf}
		\end{itemize}
	\end{itemize}
	\pause
	\linegap

	{\bf About Landslide}
	\begin{itemize}
		\item Simics module, which traces:
			\begin{itemize}
				\item Every instruction executed
				\item Every memory access read/written
			\end{itemize}
			% FIXME for next year -- how much slowdown is there? 20x?
		\item Originally supported only P3s; can now test P2s fully-automated
		\item {\em Landslide} shows how your {\em Pebbles} programs may not be stable.
	\end{itemize}
\end{frame}

@@

\begin{frame}{Landslide \& You}
	\begin{center}
	\includegraphics[width=0.85\textwidth]{landslide-new-blank.pdf}
	\end{center}
\end{frame}

\begin{frame}{Landslide \& You}
	\begin{center}
	\includegraphics[width=0.85\textwidth]{landslide-new.pdf}
	\end{center}
\end{frame}

@@

\begin{frame}{Test Suite}
	Landslide ships with 6 test cases:
	\begin{itemize}
		\item {\tt thr\_exit\_join}
		\item {\tt paraguay}
		\item {\tt rwlock\_downgrade\_read\_test}
		\item {\tt broadcast\_test} (new)
		\item {\tt paradise\_lost} (new this semester)
		\item {\tt mutex\_test} (new this semester)
	\end{itemize}
\end{frame}

%\begin{frame}{Preliminary Evaluation}
%	We tested 22 random submitted P2s from Spring 2014.
%	\begin{itemize}
%		\item Test cases:
%		\begin{itemize}
%			\item {\tt thr\_exit\_join}
%			\item {\tt paraguay}
%			\item {\tt rwlock\_downgrade\_read\_test}
%			\item {\tt broadcast\_test} (new)
%		\end{itemize}
%		\item Each test given CPU budget of 11 cores $\times$ 10 minutes.
%	\end{itemize}
%\end{frame}
%
%\begin{frame}{Preliminary Results}
%%	88 total tests were run (9,680 total CPU-minutes)
%%	\linegap
%%
%	Found 7 {\bf deterministic} initialization bugs (e.g. use-after-free)
%	\linegap
%
%	56 tests ran to completion without finding bugs:
%	\begin{itemize}
%		\item 47 exhausted the CPU budget.
%		\item 9 completed all state spaces before time ran out.
%	\end{itemize}
%	\linegap
%
%	Between 20 and 33 {\bf non-deterministic} bugs were found:
%	\begin{itemize}
%		\item 20 deadlock races among 12 P2s
%		\item 10 use-after-free races among 6 P2s
%		\item 3 misc (assert fail, segfault) among 2 P2s
%	\end{itemize}
%\end{frame}

\begin{frame}{Last Semester}
	Spring 2015: 7 groups signed up to use Landslide; 5 found bugs
	%\begin{itemize}
	%	\item (that they didn't find during stress testing)
	%\end{itemize}
	%%%% hopefully!
	\linegap

	Among all groups, 38 total tests were run
	\begin{itemize}
		\item 21 tests ran to completion without finding bugs
		\item 4 {\em deterministic} bugs
		\item 13 distinct {\em non-deterministic} bugs
			\begin{itemize}
				\item 6 in {\tt thr\_exit\_join} \small (use-after-free, deadlock, NULL, infinite loop...)
				\item 4 in {\tt paraguay} (deadlock, infinite loop)
				\item 2 in {\tt broadcast\_test} (deadlock, infinite loop)
				\item 1 in {\tt rwlock\_downgrade\_read\_test} (use-after-free)
			\end{itemize}
			%\begin{itemize}
			%	\item 6 of which required data-race PPs to expose
			%\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{A new treat since last semester}
	\texttt{mutex\_test} tries to make your mutexes fail or deadlock.
	\linegap

	Landslide will look for data races inside your mutex implementation.
	\begin{itemize}
		\item {\tt xchg}/{\tt cmpxchg}/{\tt xadd}
		\item {\tt mutex->owner = ...}
	\end{itemize}
	\linegap

	Tested 77 P2s from S'14, F'14, S'15
	\begin{itemize}
		\item 5 unrelated deterministic bugs (use-after-free, crash, etc)
			% TODO check these for false positives
			% But this was run after the arbiter fp deadlock avoidance
			% So I think it's ok
		\item 1 ``two threads got the lock at once'' race
			\begin{itemize}
				\item {\em ...that the TA who graded it didn't find!}
			\end{itemize}
		\item 7 deadlock races
		\item 4 other races (NULL crash, infinite loop, use-after-free)
	\end{itemize}
\end{frame}

@@

\begin{frame}{User Study}
	\textbf{Try Landslide on your P2!}
	\begin{itemize}
		\item Bare minimum effort: No more than 1 hour
		\begin{itemize}
			\item Clone a github URL, run setup script, run tests %, answer survey questions
			\item Landslide will automatically report test results
		\end{itemize}
		\item Full study plan: 4-8 hours of active attention
		\begin{itemize}
			\item (Estimated, including time to diagnose and fix bugs)
			\item However, many tests should run passively overnight -- start soon!
		\end{itemize}
	\end{itemize}
	\pause
	\linegap

@@

	To participate\dots
	\begin{itemize}
		\item Review this lecture and study information sheet
		\item Meet prerequisites of passing P2 tests
		\item Complete sign-up form online to get further instructions
			\begin{itemize}
				\item \url{http://tinyurl.com/landslide-p2-f15}
			\end{itemize}

		\item Use Landslide in addition to stress tests until P2 is due!
	\end{itemize}
	\linegap

@@

\begin{frame}{More Evaluation}
	\textbf{Is dynamically adding ``data race'' PPs effective?}
	\begin{itemize}
		\item 3 among preliminary bugs (9-15\%) required data races to expose.
		\item 6 among S'15 student bugs (46\%)
	\end{itemize}
	\linegap
